Ubiquitous Digital Technologies and Spatial Structure; an update

### 2021-02-01 paper/2_final_submission/tranos_ioannides.Rmd
visualisations

```{r echo=FALSE, message = FALSE, warning=FALSE}
library(stargazer)
library(tidyverse)
library(rmarkdown)
# library(data.table)
library(tidyverse)
# library(readxl)
# library(httr)
library(rprojroot)

library(ggplot2)
#library(geojsonio)
library(rgdal)
library(maptools)
library(sp)
library(sf)
library(rgeos)
#library(classInt)
#library(RColorBrewer)
library(knitr)
#library(viridis)
#library(hrbrthemes)
# library(fs)
library(betareg)
library(broom)

library(AER)
library(MASS)
#library(foreign)
library(ivpack)
# library(leaflet)
# library(rvest)
# library(xml2)
# library(tmaptools)
# library(spdep)
# library(spatialreg)

#library(acs)
#library(tidycensus)
#library(rio)
#library(wbstats)
#library(WDI)
library(kableExtra)
```

@@

Table \ref{zipf} presents the estimated $\zeta$ coefficients for the panel of countries that the second stage of the analysis focuses on using the UN urban agglomerations data.
It becomes evident that there is considerable variation in the estimates of $\zeta$ across the different countries in the sample.
Because of the empirically established heavy upper tail of data for cities and urban agglomerations, the $\zeta$ coefficient constitutes a convenient measure of dispersion.
The larger its absolute value, the thinner the upper tail; equivalently, the larger is the coefficient algebraically, the heavier the upper tail.
This key observation is basis for the second step of our methodology.

load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/UN.RData")

zipf.table <- net.model$model %>%
  dplyr::select(`as.factor(country)`, `as.factor(year)`, b) %>%
  spread(`as.factor(year)`, b) %>%
  rename(`Countries` = `as.factor(country)`)

# stargazer(zipf.table, type = "latex", digits = 3, summary = FALSE,
#           header = FALSE,
#           no.space=TRUE,
#           title = "Zipf coefficients\\label{zipf}",
#           font.size = "small",
#           column.sep.width = "1pt",
#           float.env = "sidewaystable")

kable(zipf.table, format = "latex",
      caption = "Pareto exponent\\label{zipf}",
      booktabs = T,
      digits = 2) %>%
  footnote(general = "Corrected as per Gabaix and Ibragimov (2011)") %>%
  kable_styling(latex_options = "scale_down",
                position = "left")
```

@@

stargazer(net.2sls, bb.2sls, fixed.2sls, mobile.2sls, type = "latex",
          se = list(net.2sls.rse[,"Std. Error"], bb.2sls.rse[,"Std. Error"],
                    fixed.2sls.rse[,"Std. Error"], mobile.2sls.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (3)\\label{2sls.global}",
          font.size = "small",
          column.sep.width = "1pt",
          dep.var.labels=c("Pareto exponent 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(c(rownames(summ.net.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.net.2sls$diagnostics)[2],
                             round(summ.net.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.net.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust std. errors in parenthesis",
                    "IV: Female participation in labour force",
                    "First stage regressions can be found in S2"))
```

@@

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 6
rm(list=ls())
load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/USA.RData")

@@

stargazer(model_iv1, model_iv2, type = "latex",
          se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]),
          summary = F,
          header = FALSE,
          no.space=TRUE,
          #column.separate = c(1,1,1,1),
          title = "2SLS estimation of equation (4) for USA\\label{2sls.usa}",
          font.size = "small",
          column.sep.width = "-15pt",
          align = TRUE,
          dep.var.labels=c("Difference in ranks, 2013-18"),
          covariate.labels=c("\\% of households w. internet 2013",
                             "population 2013 (log)",
                             "\\% of unemployment 2013",
                             "\\% of white population 2013",
                             "income 2013 (log)",
                             "population density 2013",
                             "\\% of employment in service 2011",
                             "commute in minutes 2013"),
          single.row = FALSE, df = FALSE,
          omit.stat = c("rsq", "f"), notes = c("Robust Std. Errors in parenthesis",
                                               "IVs: (1) Bachelors degree per hab. in 2005",
                                               "IVs: (2) Bachelors degree per hab. in 2005, Commute, minutes, 2005",
                                               "First stage regressions can be found in S2"),
          add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.fit1$diagnostics[1, "statistic"], 2),
                             round(summ.fit2$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.fit1$diagnostics)[2],
                             round(summ.fit1$diagnostics[2, "statistic"], 2),
                             round(summ.fit2$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[2, "p-value"], 2),
                             round(summ.fit2$diagnostics[2, "p-value"], 2)),
                           c(rownames(summ.fit1$diagnostics)[3],
                             round(summ.fit1$diagnostics[3, "statistic"], 2),
                             round(summ.fit2$diagnostics[3, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[3, "p-value"], 2),
                             round(summ.fit2$diagnostics[3, "p-value"], 2))
          ))

@@

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 9
rm(list=ls())
load("C:/Users/nw19521/OneDrive - University of Bristol/projects/Ioannides/ict.un.us.uk/data/data_inter/UK.RData")

@@

stargazer(model_iv1, model_iv2,
          type="latex",
          summary = F,
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (4) for the UK\\label{2sls.uk}",
          font.size = "small",
          se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]),
          dep.var.labels=c("Difference in ranks, 2011-18"),
          covariate.labels=c("download speed, 2011 (log)",
                             "population, 2011 (log)",  
                             "broadband tests per capita, 2011",
                             "\\% of unemployment, 2011",
                             "\\% of British population, 2011",
                             "population density, 2011",
                             "\\% of people working from home, 2011",
                             "employment in service, 2011 (\\%)"),
          single.row = FALSE,
          df = FALSE,
          omit.stat = c("rsq", "f"), notes = c("Robust std. errors in parenthesis", "IVs: (1) N.r of universities",
                                               "IVs: (2) N. of universities, N. of broadband tests, 2011",
                                               "First stage regressions can be found in S2"),
          add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.fit1$diagnostics[1, "statistic"], 2),
                             round(summ.fit2$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.fit1$diagnostics)[2],
                             round(summ.fit1$diagnostics[2, "statistic"], 2),
                             round(summ.fit2$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[2, "p-value"], 2),
                             round(summ.fit2$diagnostics[2, "p-value"], 2)),
                           c(rownames(summ.fit1$diagnostics)[3],
                             round(summ.fit1$diagnostics[3, "statistic"], 2),
                             round(summ.fit2$diagnostics[3, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[3, "p-value"], 2),
                             round(summ.fit2$diagnostics[3, "p-value"], 2))
          ))

>>>>

```{r echo=FALSE, message = FALSE, warning=FALSE}
library(stargazer)
library(tidyverse)
library(rmarkdown)
library(tidyverse)
library(rprojroot)

library(ggplot2)
library(rgdal)
library(maptools)
library(sp)
library(sf)
library(rgeos)
#library(classInt)
#library(RColorBrewer)
library(knitr)
#library(viridis)
#library(hrbrthemes)
# library(fs)
library(betareg)
library(broom)

library(AER)
library(MASS)
library(ivpack)
library(kableExtra)

path <- find_rstudio_root_file()
```

@@

Table \ref{zipf} presents the estimated $\zeta$ coefficients for the panel of countries that the second stage of the analysis focuses on using the UN urban agglomerations data. 
More detailed and interactive visualisations can be found in Appendix S2.
It becomes evident that there is considerable variation in the estimates of $\zeta$ across the different countries in the sample.
Because of the empirically established heavy upper tail of data for cities and urban agglomerations, the $\zeta$ coefficient constitutes a convenient measure of dispersion.
The larger its absolute value, the thinner the upper tail; equivalently, the larger is the coefficient algebraically, the heavier the upper tail.
This key observation is basis for the second step of our methodology.

```{r, echo=FALSE, warning=FALSE, out.width = "85%"}

# load data
path.data <- paste0(path, "/data/data_inter/UN.RData")
load(path.data)

# a zipf table for the appendix, which includes the Zipf coef. per year and country
zipf.table <- net.model$model %>%
  dplyr::select(`as.factor(country)`, `as.factor(year)`, b) %>%
  spread(`as.factor(year)`, b) %>%
  rename(`Countries` = `as.factor(country)`)

kable(zipf.table, format = "latex",
      caption = "Pareto exponents\\label{zipf}",
      booktabs = T,
      digits = 2) %>%
  footnote(general = "Corrected as per Gabaix and Ibragimov (2011)") %>%
  kable_styling(latex_options = "scale_down",
                position = "left")
```

@@

stargazer(net.2sls, bb.2sls, fixed.2sls, mobile.2sls, type = "latex",
          se = list(net.2sls.rse[,"Std. Error"], bb.2sls.rse[,"Std. Error"],
                    fixed.2sls.rse[,"Std. Error"], mobile.2sls.rse[,"Std. Error"]),
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (3)\\label{2sls.global}",
          font.size = "small",
          column.sep.width = "1pt",
          dep.var.labels=c("Pareto exponent 2000-18"),
          column.separate = c(1,1,1,1),
          covariate.labels=c("Internet users per 100 hab. (log)",
                             "Broadband users per 100 hab. (log)",
                             "Mobile phone users per 100 hab. (log)",
                             "Fixed phone users per 100 hab. (log)",
                             "Population density (log)",
                             "Government expenditure (\\% GDP)",
                             "Trade (\\% of GDP)",
                             "Non agriculture value added (\\% GDP)",
                             "GDP growth",
                             "GDP per capita (log)",
                             "Population (log)"),
          single.row = FALSE, df = FALSE,
          omit = c("country", "year"),
          omit.stat = c("rsq", "f"),
          add.lines = list(c(rownames(summ.net.2sls$diagnostics)[1], # use rownames as line name. in this case Weak insterumetns
                             round(summ.net.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.bb.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[1, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.net.2sls$diagnostics)[2],
                             round(summ.net.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.bb.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "statistic"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.net.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.bb.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.fixed.2sls$diagnostics[2, "p-value"], 2),
                             round(summ.mobile.2sls$diagnostics[2, "p-value"], 2)),
                           # c(rownames(summ.net.2sls$diagnostics)[3],
                           #   round(summ.net.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "statistic"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "statistic"], 2)),
                           # c("P-value",
                           #   round(summ.net.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.bb.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.fixed.2sls$diagnostics[3, "p-value"], 2),
                           #   round(summ.mobile.2sls$diagnostics[3, "p-value"], 2)),
                           c("Country fixed effects", "Yes", "Yes", "Yes", "Yes"),
                           c("Yearly fixed effects", "Yes", "Yes", "Yes", "Yes")),
          notes = c("Robust std. errors in parenthesis",
                    "IV: Female participation in labour force",
                    "First stage regressions can be found in S3"))
```

@@

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 6
rm(list=ls())

# load data
path <- find_rstudio_root_file()
path.data <- paste0(path, "/data/data_inter/USA.RData")
load(path.data)

@@

stargazer(model_iv1, model_iv2, type = "latex",
          se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]),
          summary = F,
          header = FALSE,
          no.space=TRUE,
          #column.separate = c(1,1,1,1),
          title = "2SLS estimation of equation (4) for USA\\label{2sls.usa}",
          font.size = "small",
          column.sep.width = "-15pt",
          align = TRUE,
          dep.var.labels=c("Difference in ranks, 2013-18"),
          covariate.labels=c("\\% of households w. internet 2013",
                             "population 2013 (log)",
                             "\\% of unemployment 2013",
                             "\\% of white population 2013",
                             "income 2013 (log)",
                             "population density 2013",
                             "\\% of employment in service 2011",
                             "commute in minutes 2013"),
          single.row = FALSE, df = FALSE,
          omit.stat = c("rsq", "f"), notes = c("Robust Std. Errors in parenthesis",
                                               "IVs: (1) Bachelors degree per hab. in 2005",
                                               "IVs: (2) Bachelors degree per hab. in 2005, Commute, minutes, 2005",
                                               "First stage regressions can be found in S3"),
          add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.fit1$diagnostics[1, "statistic"], 2),
                             round(summ.fit2$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.fit1$diagnostics)[2],
                             round(summ.fit1$diagnostics[2, "statistic"], 2),
                             round(summ.fit2$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[2, "p-value"], 2),
                             round(summ.fit2$diagnostics[2, "p-value"], 2)),
                           c(rownames(summ.fit1$diagnostics)[3],
                             round(summ.fit1$diagnostics[3, "statistic"], 2),
                             round(summ.fit2$diagnostics[3, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[3, "p-value"], 2),
                             round(summ.fit2$diagnostics[3, "p-value"], 2))
          ))

@@

```{r, echo=FALSE, warning=FALSE, results='asis'}
# table 9
rm(list=ls())

path <- find_rstudio_root_file()
path.data <- paste0(path, "/data/data_inter/UK.RData")
load(path.data)

@@

stargazer(model_iv1, model_iv2,
          type="latex",
          summary = F,
          header = FALSE,
          no.space=TRUE,
          title = "2SLS estimation of equation (4) for the UK\\label{2sls.uk}",
          font.size = "small",
          se = list(rob.model_iv1[,"Std. Error"], rob.model_iv2[,"Std. Error"]),
          dep.var.labels=c("Difference in ranks, 2011-18"),
          covariate.labels=c("download speed, 2011 (log)",
                             "population, 2011 (log)",  
                             "broadband tests per capita, 2011",
                             "\\% of unemployment, 2011",
                             "\\% of British population, 2011",
                             "population density, 2011",
                             "\\% of people working from home, 2011",
                             "employment in service, 2011 (\\%)"),
          single.row = FALSE,
          df = FALSE,
          omit.stat = c("rsq", "f"), notes = c("Robust std. errors in parenthesis", "IVs: (1) N.r of universities",
                                               "IVs: (2) N. of universities, N. of broadband tests, 2011",
                                               "First stage regressions can be found in S3"),
          add.lines = list(c(rownames(summ.fit1$diagnostics)[1],                    # use rownames as line name. in this case Weak insterumetns
                             round(summ.fit1$diagnostics[1, "statistic"], 2),
                             round(summ.fit2$diagnostics[1, "statistic"], 2)),
                           c(rownames(summ.fit1$diagnostics)[2],
                             round(summ.fit1$diagnostics[2, "statistic"], 2),
                             round(summ.fit2$diagnostics[2, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[2, "p-value"], 2),
                             round(summ.fit2$diagnostics[2, "p-value"], 2)),
                           c(rownames(summ.fit1$diagnostics)[3],
                             round(summ.fit1$diagnostics[3, "statistic"], 2),
                             round(summ.fit2$diagnostics[3, "statistic"], 2)),
                           c("P-value",
                             round(summ.fit1$diagnostics[3, "p-value"], 2),
                             round(summ.fit2$diagnostics[3, "p-value"], 2))
          ))
